### 2012-10-27 2.tex
Improvements to the appendices, primarily.

%\citet{Fischer-Jorgensen1952}
%\citet{Saporta1958}
%\citet{Saporta1957}

\citet{Halle1962} and \citet{Chomsky1965} use two nonce words, \emph{blick} [blɪk] and \emph{bnick} [bnɪk], to illustrate that speakers internalize generalizations about possible and impossible words in their language. While \emph{blick} is unattested, but in some sense a ``possible'' word of English, \citeauthor{Chomsky1965} observe that speakers immediately recognize \emph{bnick} to be ill-formed, an impossible word, though it is well-formed in other languages (e.g., Moroccan Arabic \emph{bniqa} `closet').

@@

\begin{quote}
A defect of current grammatical acounts of phonotactics is that they render simple up-or-down decisions concerning well-formedness and cannot account for gradient judgements. But when judgements are elicited in a controlled fashion from speakers, they always emerge as gradient, including all intermediate values. \citep[371]{Shademan2006} 
\end{quote}

@@

\begin{table}
\label{counts}
\centering
\begin{tabular}{l r r r}
\toprule
        & subjects & items & observations \\
\midrule
Scholes & 33       & 66    & 2,079        \\
A\&H    & 24       & 88    & 2,112        \\
\bottomrule
\end{tabular}
\caption{Idk.}
\end{table}

@@

>>>>

\citet{Halle1962} and \citet{Chomsky1965} use the nonce words \emph{blick} [blɪk] and \emph{bnick} [bnɪk] to illustrate the contrast between possible words, accidental gaps in the lexicon, and impossible words, structurally excluded words, respectively. In \emph{SPE}, \citealt{SPE} introduce a third nonce word, \emph{bnzk}, so to illustrate that speakers have intuitions about multiple degrees of wellformedness. Both \emph{bnick} and \emph{bnzk} are possible words in other languages: stop-nasal onsets are found in Moroccan Arabic (e.g., \emph{bniqa} `closet') and stop-nasal-fricative-stop words in Imdlawn Tashlhiyt Berber \citep{Dell1984}. While neither are permitted in English, there is some real sense in which \emph{bnzk} is even less English-like than \emph{bnick}. \citeauthor{SPE} write:

In recent years there has been increased interest in the ``problem of admissibility'', and particularly in the multiple degrees of wellformedness inferred from wordlikeness tasks:

\begin{quote}
When native speakers are asked to judge made-up (nonce) words, their intuitions are rarely all-or-nothing. In the usual case, novel items fall along a gradient cline of acceptability. \citep[][9]{Albright2009a}

In the particular domain of phonotactics gradient intuitions are pervasive: they have been found in every experiment that allowed participants to rate forms on a scale.
\citep[][382]{Hayes2008a}

\ldots{}when judgements are elicited in a controlled fashion from speakers, they always emerge as gradient, including all intermediate values. \citep[371]{Shademan2006} 
\end{quote}

%A defect of current grammatical acounts of phonotactics is that they render simple up-or-down decisions concerning well-formedness and cannot account for gradient judgements. 
% shademan

There has been increased interest in the gradient aspects of wordlikeness judgements in recent years.

@@

@@

\begin{table}
\label{counts}
\centering
\begin{tabular}{l r r r}
\toprule
        & subjects & items & observations \\
\midrule

\begin{table}
\centering
\begin{tabular}{l r r r}
\toprule
                        & subjects & items & trials \\
\midrule
\citealt{Albright2003a} & 24       & 87    & 2,088  \\
\citealt{Albright2007}  & 68       & 40    & 2,720  \\
\citealt{HayesInPress}  & 29       & 80    & 2,320  \\
\citealt{Scholes1966}   & 33       & 63    & 2,178  \\
\midrule
\textsc{Total}          & 154      & 270   & 9,306  \\
\bottomrule
\end{tabular}
\caption{Subject and item counts}
\label{counts}
\end{table}

\bottomrule
\end{tabular}
\caption{Idk.}
\end{table}

@@

% other ratings, but por qué?:
% 
% Hayes2000
% 
% Albright2003a
% Albright2003b
% Prasada1993 
% 
% Bard1996

### 2012-10-27 3.tex
Improvements to the appendices, primarily.

\noindent 
In the full specification analysis adopted here, reharmonization is the result of \textsc{Backness Harmony} applying within the reduplicant. On the other hand, the lack of reharmonization in the reduplicants of disharmonic roots suggests that the [$-$\textsc{Backness Harmony}] exception feature is copied under reduplication.\footnote{The strongest argument that reduplication preserves diacritic features of the root would be a case where lexical exceptionality surfaces only in reduplicants. Kinande reduplication, documented by \citet{Mutaka1990} and further discussed by \citet{Downing2000}, provides such a case. Verb reduplication requires a bisyllabic reduplicant. This clearly has synchronic force in the gramar, since reduplicated monosyllabic roots actually have three copies of the root, and since reduplicated forms of many trisyllable verbs are ineffable. However, a few trisyllabic verbs exceptionally show full reduplication. Exceptionality appears to be a lexical property (since it is not predictable), but it surfaces only in the reduplicant.} \citet{Silverman2000} notes that static phonotactic generalizations are not preserved by operations like reduplication and truncation: under this definition, root harmony is not static.

@@

\citet{Becker2011} attribute the inactivity of \textsc{Labial Attraction} in wordlikeness to a claim that it is phonologically unnatural, and therefore goes unlearned. This claim is difficult to evaluate insofar as phonotactic naturalness has continually eluded formalization and \citeauthor{Becker2011} neither propose nor refer to any theory thereof. \citeauthor{Becker2011} use the lack of phonetic precursors as a diagnostic of unnaturalness; however, there is extensive evidence that ``unnatural'' generalizations are in fact learned \citep[e.g.,][]{Anderson1981,Bach1972,Blevins2004,Buckley2000a,Hayes2009,Pierrehumbert2006c}. Furthermore, other functionally-oriented theorists have considered \textsc{Labial Attraction} sufficiently natural \citep[e.g.,][]{NiChiosain1993,Ito1993,Ito1995b}.

It is not even obvious that it is desirable to exclude \textsc{Labial Attraction} from the grammar. \citet[394, fn. 2]{Inkelas1997} suggest that \textsc{Labial Attraction} may have even induced alternations at one point in the history of Turkish. Beyond Turkish, the rounding of a high back vowel after a labial consonant, as in the alternative formulation of \textsc{Labial Attraction} proposed by \citet{Inkelas2001}, is widely attested \citep[e.g.,][]{Vaux1993}. It is also phonetically natural, as both labial consonants and high round vowels are distinguished by a very low first and second formant. As shown in Table \ref{lasi}, it is even a statistically reliable generalization about the Turkish lexicon. However, it is not reflected in alternations or in wordlikeness judgements.

>>>>

\noindent 
In the full specification analysis adopted here, reharmonization is the result of \textsc{Backness Harmony} applying within the reduplicant. On the other hand, the lack of reharmonization in the reduplicants of disharmonic roots suggests that the [$-$\textsc{Backness Harmony}] exception feature is copied under reduplication.\footnote{There is reason to believe that reduplicants bear lexical diacritics. Kinande reduplication, documented by \citet{Mutaka1990} and further discussed by \citet{Downing2000}, provides such a case. Verb reduplication requires a bisyllabic reduplicant. This clearly has synchronic force in the gramar, since reduplicated monosyllabic roots actually have three copies of the root, and since reduplicated forms of many trisyllable verbs are ineffable. However, a few trisyllabic verbs exceptionally show full reduplication. Exceptionality appears to be a lexical property (since it is specific to certain lexical entries), but it surfaces only in the reduplicant.} \citet{Silverman2000} notes that static phonotactic generalizations are not preserved by operations like reduplication and truncation: under this definition, root harmony is not static.

@@

\citet{Becker2011} attribute the inactivity of \textsc{Labial Attraction} in wordlikeness to a claim that it is phonologically unnatural, and therefore goes unlearned. This claim is difficult to evaluate insofar as phonotactic naturalness has continually eluded formalization and \citeauthor{Becker2011} neither propose nor refer to any theory thereof. \citeauthor{Becker2011} use the lack of phonetic precursors as a diagnostic of unnaturalness; however, there is extensive formal and experimental evidence for the learning of ``unnatural'' generalizations \citep[e.g.,][]{Anderson1981,Bach1972,Blevins2003,Buckley2000a,Hayes2009,Pierrehumbert2006c,Seidl2005}. Furthermore, other functionally-oriented theorists have considered \textsc{Labial Attraction} sufficiently natural \citep[e.g.,][]{NiChiosain1993,Ito1993,Ito1995b}.

It is not even obvious that it is desirable to exclude \textsc{Labial Attraction} as a possible rule. \citet[394, fn. 2]{Inkelas1997} suggest that \textsc{Labial Attraction} may have even induced alternations at one point in the history of Turkish. Beyond Turkish, the rounding of a high back vowel after a labial consonant, as in the alternative formulation of \textsc{Labial Attraction} proposed by \citet{Inkelas2001}, is widely attested \citep[e.g.,][]{Vaux1993}. It is also phonetically natural, as both labial consonants and high round vowels are distinguished by a very low first and second formant. As shown in Table \ref{lasi}, it is even a statistically reliable generalization about the Turkish lexicon. However, it is not reflected in alternations or in wordlikeness judgements.

### 2012-10-27 4.tex
Improvements to the appendices, primarily.

Early work in generative phonology recognized that neutralizing, surface-true phonological alternations impose constraints on surface forms and underlying representations (\citealp[283]{Anderson1974}, \citealp[382]{SPE}, \citealp[205f.]{Dell1973}, \citealp[22f.]{SPR}, \citealp{Kenstowicz1977}: chapter 3, \citealp[28f.]{Stampe1973}, \citealp[410f.]{Stanley1967}). Despite isolated claims to the contrary (e.g., \citealp[297]{Hale1965}; \citealp[212f.]{Postal1968}), however, the \emph{communis opinio} (including many of the authors cited above) holds that not all phonotactics---that is, language-specific sound sequence co-occurrence restrictions---can be derived in this fashion, and phonotactic knowledge is independent of the phonology.

@@

A list of medial clusters is gathered by applying a number of filters and transformations to an English lexical database. The resulting list of clusters is reproduced in Appendix \ref{appendixB} and in electronic form at the author's website.

@@

\begin{example}[Lexical effects of \textsc{Obstruent Voice Assimilation}] \label{ovatable}
\begin{tabular}{l r r r r}
\toprule
           & attested & unattested & \% saturation & $p$-value                   \\
\midrule
conforming & 80       & 370        & 18          & \multirow{2}{*}{1.1\e{-11}} \\
violating  &  6       & 264        & 2           \\
\bottomrule
\end{tabular}
\end{example}

@@

\begin{table} \centering
\begin{tabular}{l | r r r r}
\toprule
                     & accuracy & precision & recall & $F_1$ \\
\midrule
Null baseline        & 0.813    & n.a.      & n.a.   & n.a.  \\
Derived constraints  & 0.857    & 0.401     & 0.708  & 0.512 \\
Expected frequency   & 0.839    & 0.210     & 0.750  & 0.328 \\
\citet{Hayes2008a}   & 0.877    & 0.777     & 0.642  & 0.703 \\
\bottomrule
\end{tabular}
\caption{The \citeauthor{Hayes2008a} model makes the most accurate predictions about possible and impossible syllable contact clusters, though it is only slightly better than the derived constraint baseline.}
\label{cmresults}
\end{table}

@@

18.8\% of the 840 possible clusters are attested. Consequently, a null baseline which predicts no clusters to be attested trivially achieves 81.2\% accuracy. The other metrics are undefined for this baseline.

@@

The three derived constraints define a structured baseline, in which attestation is the predicted outcome for all clusters which not violate the \emph{SPE}-derived generalizations, and non-attestation otherwise. This produces a significant improvement to the null baseline (accuracy = 0.857; sign test $p = 2.16$\e{-05}). The high recall (0.708) but low precision (0.401) of this model indicates that there are more false positives (i.e., unattested clusters which do not violate any constraint) than false negatives (i.e., exceptions to the three rules). It is not clear that low precision is a flaw, since some of these false positives might be accidental gaps---that is, clusters that are well-formed, but unattested.

>>>>

Early work in generative phonology recognized that neutralizing, surface-true phonological alternations impose constraints both on surface forms and underlying representations (\citealp[283]{Anderson1974}, \citealp[382]{SPE}, \citealp[205f.]{Dell1973}, \citealp[22f.]{SPR}, \citealp{Kenstowicz1977}: chapter 3, \citealp[28f.]{Stampe1973}, \citealp[410f.]{Stanley1967}).

In the literature there are both isolated claims that 
There are both isolated claims that

Despite isolated claims to the contrary (e.g., \citealp[297]{Hale1965}; \citealp[212f.]{Postal1968}),

\citet{Kisseberth1970b}

however, the \emph{communis opinio} still is that that not all phonotactics---that is, language-specific sound sequence co-occurrence restrictions---can be derived in this fashion, and phonotactic knowledge is independent of the phonology.

%FIXME
% \citet[255]{Anderson1974}, discussing this fact, writes that \citeauthor{Haugen} ``gives an argument that shows the necessity of including syllables as units in underlying representations''.

%FIXME
(for dissenting views, see \citealt{Blevins2003} and \citealt{Steriade1999})

Linguists have long recognized underlying representations as privileged domains for stating phonotactic generalizations (e.g., \citealt{Bloomfield1930,Jakobson1932,Pike1947b}).

More recently, a similar role for the syllable has been recognized 
(e.g., \citet{Haugen1956}, \citet{Hooper1973}, \citealt{Kahn1976})

@@

A list of medial clusters is gathered by applying a number of filters and transformations to an English lexical database. The resulting list of clusters is reproduced in Appendix \ref{appendixC} and in electronic form at the author's website.

@@

\begin{example}[Lexical effects of \textsc{Obstruent Voice Assimilation}] \label{ovatable}
\begin{tabular}{l r r r r}
\toprule
           & attested & unattested & \% saturation & $p$-value                   \\
\midrule
conforming & 80       & 370        & 18\%          & \multirow{2}{*}{1.1\e{-11}} \\
violating  &  6       & 264        & 2\%           \\
\bottomrule
\end{tabular}
\end{example}

@@

\begin{table} \centering
\begin{tabular}{l | r r r r | r}
\toprule
                     & accuracy & precision & recall & $F_1$ \\
\midrule
Null baseline        &     .808 &      .808 &  1.000 &  .894 \\
\citet{Hayes2008a}   &     .837 &      .948 &   .852 &  .898 \\ %& .002       \\
Expected frequency   &     .835 &      .845 &   .983 &  .909 \\ %& 1.1\e{-05} \\
Derived constraints  &     .840 &      .922 &   .885 &  .903 \\
DC and EF            &     .864 &      .871 &   .989 &  .926 \\ %& 1.9\e{-13} \\
\bottomrule
\end{tabular}
\caption{FIXME}
\label{cmresults}
\end{table}

@@

19\% of the 880 possible clusters are attested. Consequently, a null baseline which predicts that no clusters are possible will be 81\% accurate, trivially. Because this baseline FIXME RECALL FIXME PRECISION

@@

The three derived constraints define a structured baseline, in which attestation is the predicted outcome for all clusters which not violate the \emph{SPE}-derived generalizations, and non-attestation otherwise. This results in a significant improvement to the null baseline (sign test $p = 0.006$).  This model has poor recall, however, because many well-formed clusters are unattested. This is not a flaw, but rather the signature of accidental gaps.

### 2012-10-27 pre.tex
Improvements to the appendices, primarily.

This negative heuristic also allows for the possibility of using psycholinguistic evidence to decide between competing phonological analyses. For instance, there is a long-standing debate concerning the representation of Sanskrit aspiration alternations like those represented by \emph{bodhati}-\emph{bhotsyati} `he wakes-he will wake'. One analysis, going back to the ancient grammarian Pāṇini \citep{Sag1974} and to \citet[\S141f.]{Whitney1889}, posits an underlying representation /bud\asp{}-/ with aspiration shifting leftward in certain contexts: modern proponents of this analysis include \citet{Borowsky1983}, \citet[][59f.]{Hoenigswald1965}, \citet{Kaye1985}, \citet{Sag1974,Sag1976} \citet{Schindler1976}, \citet{Stemberger1980}, and \citet[][\S141f.]{Whitney1889}. An alternative analysis posits an underlying representation /b\asp{}ud\asp{}-/ and an apsirate dissimilation process (\citealt{Anderson1970}, \citealt{Hoard1975}, \citealt[\S3.2]{Kiparsky1965}, \citealt{Phelps1973}, \citealt{Phelps1975b}, \citealt[109f.]{Zwicky1965}). It is not my intention to adjudicate between these competing analyses, or to evaluate the claim that both accounts are deeply flawed \citep{Janda1989}, but simply to note that they make different predictions under the theory proposed here. While the former, aspirate shift analysis denies the existence of diaspirate roots, this fact lacks synchronic motivation, so the co-occurrence restriction go unlearned. On the other hand, under the aspirate dissimilation account, underlying diaspirate roots exist but speakers internalize synchronic processes responsible for their absence on the surface.

>>>>

This negative heuristic also allows for the possibility of using psycholinguistic evidence to decide between competing phonological analyses. For instance, there is a long-standing debate concerning the representation of Sanskrit aspiration alternations like those represented by \emph{bodhati}-\emph{bhotsyati} `he wakes-he will wake'. One analysis, going back to the ancient grammarian Pāṇini \citep{Sag1974} and to \citet[\S141f.]{Whitney1889}, posits an underlying representation /bud\asp{}-/ with aspiration shifting leftward in certain contexts: modern proponents of this analysis include \citet{Borowsky1983}, \citet[][59f.]{Hoenigswald1965}, \citet{Kaye1985}, \citet{Sag1974,Sag1976} \citet{Schindler1976}, \citet{Stemberger1980}, and \citet[][\S141f.]{Whitney1889}. An alternative analysis posits an underlying representation /b\asp{}ud\asp{}-/ and an aspirate dissimilation process (\citealt{Anderson1970}, \citeyear[46f.]{Anderson1974}, \citealt{Hoard1975}, \citealt[\S3.2]{Kiparsky1965}, \citealt{Phelps1973}, \citealt{Phelps1975b}, \citealt[109f.]{Zwicky1965}). It is not my intention to adjudicate between these competing analyses, or to evaluate the claim that both accounts are deeply flawed \citep{Janda1989}, but simply to note that they make different predictions under the theory proposed here. While the former, aspirate shift analysis denies the existence of diaspirate roots, this fact lacks synchronic motivation, so the co-occurrence restriction go unlearned. On the other hand, under the aspirate dissimilation account, underlying diaspirate roots exist but speakers internalize synchronic processes responsible for their absence on the surface.
